Depression and suicidality as credible signals of need

### 2021-05-17 Paper.rmd
Getting close to final version

@@

<!-- Our results suggest that the costs of depression and suicidality, which in mainstream theories are taken as evidence of brain dysfunction, might instead function to help victims elicit support from reluctant social partners with whom they are in conflict. -->

@@

The main limitation of this study is that is has less ecological validity than real-world observations of depressed individuals interacting with their social partners. It is almost certain that participants experienced this study differently than they would if the event actually happened to them. This may have biased results in a pro-signaler direction if the lack of real costs of helping made support feel less costly or if there was a social desirability bias toward helping [@grimm2010]. It may have also biased participants against helping if they could not fully imagine the characters in the story as kin or important cooperative partners, while the survey's short duration may have weakened the strength of the costlier signals as bargaining tools.

Another limitation is that our control included a signal in the form of a verbal request, preventing comparisons to non-signaling strategies. Similarly, our design did not include vignettes with male signalers. For this reason, we have no data on the possibility of sex differences in the effectiveness of the signaling strategies examined. Although not predicted theoretically, such differences are possible if the costs of signaling vary between the sexes due to differential access to alternative bargaining strategies [@HagenRosenstrom2016] or if one sex tends to suffer greater negative reputational effects  when displaying the emotions and behaviors in the vignettes. It is also possible the costliness of the situations presented in the vignettes differ by sex.

In terms of data quality our largest limitation is that, compared to the US sample, far more Indian participants were filtered out by our attention checks, something which could be consistent with botting, unfamiliarity with English, or low effort responses [@kennedy2020]. Assuming this corresponded with lower-quality responses among those who passed the attention checks, the weak signal effect in the thwarted marriage vignette may simply be due to added noise rather than differences in the scenario or the effectiveness of the signals compared to those in the US. Another concern relevant to all vignettes is that our decision to anchor the T1 sliders at 0 may have resulted in participants being more likely to report extreme values.

<!-- Slider anchor  -->
<!-- Regex: stopped in SI -->
<!-- T\d \w+ -->
# Conclusion

Our results suggest that the costs of depression and suicidality, which in mainstream theories are taken as evidence of brain dysfunction [@SymeHagen2021], might instead function to help victims elicit support from reluctant social partners with whom they are in conflict. This coupled with findings that depression can simultaneously elicit both positive and negative responses [@HokansonEtAl1986; @stephens1987], suggests that more attention should be paid to the potential for positive responses towards depressed individuals, especially from individuals who normally receive substantial benefits from the depressed or suicidal individual.

>>>>

Some evolutionary theorists have questioned the bargaining model based on evidence that depression is aversive to social partners [@nettle2004; @allen2006; @nesse2006; @myers2016]. These critics fail to acknowledge the central role of *conflict* in the signaling/bargaining models. Anger causes negative reactions in targeted social partners, yet it is probably an adaptation that exploits advantages in physical or social formidability to force concessions from others [@sell2009]. Aversive responses to bargaining are expected throughout the process, with potential helpers likely signaling the costliness of providing support in turn. Importantly, such responses do not prevent the delivery of meaningful benefits [@SymeHagen2021].

@@

@@

The main limitation of this study is that is has less ecological validity than real-world observations of depressed individuals interacting with their social partners. This may have biased results in a pro-signaler direction if the lack of real costs of helping made support feel less costly or if there was a social desirability bias toward helping [@grimm2010]. It may have also biased participants against helping if they could not fully imagine the characters in the story as kin or important cooperative partners, and the survey's short duration may have weakened the strength of the costlier signals as bargaining tools.

Our design did not include vignettes with male signalers. For this reason, we have no data on the possibility of sex differences in the effectiveness of the signaling strategies examined. Although not predicted theoretically, such differences are possible if the costs of signaling vary between the sexes due to differential access to alternative bargaining strategies [@HagenRosenstrom2016] or if one sex tends to suffer greater negative reputational effects  when displaying the emotions and behaviors in the vignettes. It is also possible the costliness of the situations presented in the vignettes differ by sex.

Compared to the US sample, far more Indian participants failed our attention checks, which is consistent with botting, unfamiliarity with English, or low-effort responses [@kennedy2020]. If this high failure rate indicates lower-quality responses among those who passed the attention checks, the weak signal effect in the thwarted marriage vignette may simply be due to greater noise rather than differences in the scenario or the effectiveness of the signals compared to those in the US. Another concern relevant to all vignettes is that our decision to anchor the T1 sliders at 0 may have resulted in participants being more likely to report extreme values.

<!-- Slider anchor  -->
<!-- Regex: stopped in SI -->
<!-- T\d \w+ -->

# Conclusion

Depression is costly, sometimes resulting in death by suicide. Our results suggest that these costs, which mainstream theories take as evidence of brain dysfunction [@SymeHagen2021], function to help victims of adversity elicit support when their true level of need is private information and they have conflicts with social partners. These findings align with real-world evidence that depression simultaneously elicits positive and negative responses from social partners [@SymeHagen2021]. If the bargaining model is correct, depression is an important social problem, not a medical one [@syme2020].

<!-- This coupled with , suggests that more attention should be paid to the potential for positive responses towards depressed individuals, especially from social partners who receive substantial benefits from the depressed or suicidal individual. -->
